
\documentclass{report}

\input{../../preamble}
\input{../../macros}
\input{../../letterfonts}

\title{INDENG 164: Assignment 3}
\author{Ryan Lin}
\date{}

\begin{document}
\maketitle
\section{Question 1}

A captain is transporting three items across a river: a wolf (W), a goat (G), and a cabbage (C).
If the captain is not present on a bank, then:

\begin{itemize}
	\item  the wolf and goat cannot be left together on that bank; and
	\item the goat and cabbage cannot be left together on that bank.
\end{itemize}
Using binary decision variables of your choice, write constraint(s) that enforce these rules.


Let $ Y_{i} = \begin{cases}
	1, & \text{if the captain is present on bank $ i $ } \\ 
	0, & \text{if the captain is not present on bank $ i $ }
\end{cases} $ 

Let $ W_{i}, G_{i}, C_{i} $ be if 1 the wolf, goat, and cabbage are on bank $ i $ respectively, otherwise 0. 

\begin{align*}
	W_{i} + G_{i} \le 1 + Y_{i} \\
	G_{i} + C_{i} \le 1 + Y_{i} \\
	Y_i, W_{i}, G_{i}, C_{i} \in \{0,1\}
\end{align*}

\newpage
\section{Question 2} 

Let $ x_{ij} $ defined for each district $ i $ and each county $ j $, where $ x_{ij} \begin{cases}
	1, \text{if county $ j $ is assigned to district $ i $ } \\
	0, \text{otherwise}
\end{cases}$

Let $ D_{i} $ difference between Democratic and Republican votes in district $ i $ 

Let $ Y_{j} $ be the difference between Democratic and Republican votes in county $ j $ 


Maximize the number of votes that the Democratic party wins district 2 by. If we can get the Democratic party to win district 2 by a margin of at least 100 votes while still winning districts 1 and 3 by a margin of at least 100 votes in each district, the Democratic party will win all three districts. 

Constraints: 
\begin{itemize}
	\item Each count to be assigned to exactly 1 district \\
	\item Each district to consist of at least one county \\
	\item Democratic party to still win districts 1 and 3 by a margin of at least 100 votes
\end{itemize}


\begin{align*}
	\max_{}: \quad & D_{2}  \\
\text{s.t}: \quad &  \sum_{i=1}^{3} x_{ij} = 1 \forall j\\
& \sum_{j=1}^{33} x_{ij} \ge 1 \forall i\\ 
& D_{i} = \sum_{j=1}^{33} Y_{j} x_{ij} \forall i \\
& D_{1}  \ge 100 \\
& D_{2} \ge 100\\ 
& D_{3}  \ge 100 \\
& x_{ij} \in \{0,1\} \forall i,j
\end{align*}

Constraint 1 forces each county to only be assigned to 1 district. Constraint 2 forces each district to have at least 1 county. Constraint 3 sets the difference of a district's votes to be the sum of the individual counties in the district. Constraints 4-6 ensures we still win all districts by at least 100 votes.  

Optimal Solution: $ \boxed{76192}$ votes 


\newpage
\section{Question 3}


Decision Variables:
\begin{itemize}
	\item $ X_{ijk} = \begin{cases}
			1 &\text{if (i,j) entry is a number k } \\
			0 & \text{otherwise}
	\end{cases} $ 
\end{itemize}
Where $ (i,j) $ denotes row $ i $ from top to bottom and column $ j $ from left to right. 

We start with: 


\begin{align*}
\min_{}: \quad & 0 \\
\text{s.t}: \quad & \sum_{i=1}^{9} x_{i,j,k} = 1 \forall j, k \\ % all rows have only 1 instance of a num
		  & \sum_{j=1}^{9} x_{i,j,k} = 1 \forall i,k \\% all cols have only 1 instance of a num
		  & \sum_{i=x}^{x+2} \sum_{j=y}^{y+2} x_{ijk} = 1\forall k, \forall x,y \in \{3,6,9\} \\
		  & \sum_{k=1}^{9} x_{ijk} = 1 \forall i,j\\ % all cells only have 1 num
		  & \sum_{k=1}^{9} x_{ijk} = 1 \forall i,j\\
		  & x_{ijk} \in \{0,1\} \forall i,j,k\\ 
	&x_{1,6,1} = 1 \\ 
	&x_{2,1,2} = 1, 
	x_{2,2,7} = 1,
	x_{2,5,9} = 1, 
	x_{2,7,5} = 1, \\
	&x_{3,2,8} = 1, 
	x_{3,6,5} = 1,
	x_{3,9,3} = 1\\ 
	&x_{4,3,8} = 1, 
	x_{4,6,3} = 1,
	x_{4,8,2} = 1 \\
	&x_{5,2,5} = 1,
	x_{5,4,1} = 1, 
	x_{5,6,2} = 1,
	x_{5,8,9} = 1\\
	&x_{6,2,1} = 1,
	x_{6,5,5} = 1,
	x_{6,7,7} = 1\\
	&x_{7,1,5} = 1,
	x_{7,4,6} = 1,
	x_{7,8,3} = 1\\
	&x_{8,3,9} = 1,
	x_{8,5,1} = 1,
	x_{8,8,6} = 1,
	x_{8,9,2} = 1\\
	&x_{9,4,2} = 1\\
\end{align*}

\end{document}
