
\documentclass{report}

\input{../../preamble}
\input{../../macros}
\input{../../letterfonts}

\title{\Huge{IEOR 173}\ Lecture Notes}
\author{\huge{Ryan Lin}}
\date{}

\begin{document}
\maketitle
\newpage% or \cleardoublepage
% \pdfbookmark[<level>]{<title>}{<dest>}
\pdfbookmark[section]{\contentsname}{too}
\tableofcontents
\pagebreak

\chapter{}
\section{Lecture 03}
\subsection{Joint Distributions}
\[P(x,y) = P(X \le x, y \le y)\]
This is a cumulative distribution for joint distributions. 

If $X \perp Y$ (independent), (perpendicular), thee the join distribution is the product of the marginals. $P(X,Y) = P(X)P(Y) \forall X, Y$ 

\subsection{Cov + Correlations} 
Covariance is a measure of linear dependence.
\[Cov(X,Y) = E[(x-EX)(Y-EY)] = EXY - EXEY\]

Interpretation: If they vary together, Cov is larger. 

Properties
\begin{enumerate}
	\item Cov(X,X) = Var(x)
	\item Cov(X,Y) = Cov(Y,X)
	\item Cov(cX, Y) = cCov(X,Y)
	\item Cov(X, Y+z) = Cov(X,Y) + Cov(X,Z)
\end{enumerate}

Correlation:
\[p(X,Y) = \frac{Cov(X,Y)}{sd(X) \cdot sd(Y)}\]
If 
\begin{itemize}
	\item  > 0, pos correlation
	\item < 0, negative correlation
	\item = 0 No correlation
\end{itemize}

\nt{Independent does imply uncorrelated, however uncorrelated does not imply independent.

	Ex: If $P(-1,1) = P(0,0) = P(1,1) = \frac{q}{3}$. then $E[XY] = \frac{0}{3} = 0$. However, $E[X] = 0, EY = ?$, therefore $E[XY] = 0$ and $Cov(X,Y) = 0$. There is no linear relationship, however there is a dependency for $y=x^2$. 

}

Therefore, we can generalize our variance formula

\[Var(A+\sum b_i X_i) =\sum b_i^2 Var(X_i) + \sum_{\forall ij, i\neq j} b_ib_j Cov(X_i, X_j) \]

\subsection{Hat Example}
\ex{Lists of number}{Imagine go to party, men take off hats, women keep on hats. Each man randomly takes a hat and leaves. 

Let $X$ = \# of matches in $N$ pairs (man and his hat)

What is EX and Var(X)? 

If two men, the only possibility of matches are 0 and 2. The mean is 1. 
If three men, a similar pattern occurs. 

\[
	EX = 1 \forall N
\]
As $N$ grows larger, the probability decreases, but there is more men to balance it out. 

\[Var(X) = 1 \forall N \ge 2\]

We can prove this by using Indicators. 


Let $X = \sum^N I_i$, where $I_i = I{\text{it is a match}} \sim Bern(\frac{1}{N})$

\[EX = E[\sum I] = \sum^N E[I] = N \frac{1}{N} = 1\]

\[Var(I_i) = E[I^2] - [EI]^2 = p - p^2 = p(1-p) = \frac{1}{N}- \frac{1}{N^2} = \frac{N-1}{N^2}\]
\[Cov(I_i, I_j) = E[I_iI_j] = EI_i EI_j\]
\[E[I_i I_j] = P(\text{both i and j match}) = P(I-i = I_j = 1)\]
\[P(I_i = 1)P(I_j =1 \mid I_i = 1) = \frac{1}{N} \cdot \frac{1}{N-1}\]
Conceptually, we can think of event $j$ following event $j$

\[Cov(I_i, I_j) = \frac{1}{n(n-1)} - \left( \frac{1}{n} \right)^2 = \frac{1}{N^2(N-1)}\]

\[Var(X) = Var(\sum I_i) = \sum Var(I) + \sum Cov(I_i, I_j) = N \cdot \frac{N+1}{N^2} + N(N-1) \frac{1}{N^2(N-2)} = 1\]

The calculation will be left as an exercise to the future reader.


As $N \rightarrow \infty$,

\[
	Cov(I_i, I_j) \rightarrow 0
\]

\[
	Var(I_i) \rightarrow \frac{1}{N}
\]
\[
	X \rightarrow Bin(N, \frac{1}{N}) \sim Poisson(1)
\]

Intuitively, if there are thousands of men there, whether or not 1 person gets his hat is not really going to affect the chances another gets his.  
}
\subsection{Start condition (Ch 3)} 
What if we have dependent random variables and want to compute conditional probabilities? 

Conditions:
\begin{enumerate}
	\item discrete: $P(X,y) = P(X=x, X=y), P(x,y) = P(X\le x, Y\le y) = \sum_{\le x}\sum_{\le y} P(x,y)$
	\item continuous: $f(x,y) = F(x,y) = \int_{-\infty}^x \int_{-\infty}^y f(u,v) dudv$
	\item marginal: $F_X(x) = P(X\le x) = F(X,\infty)$ 
	\item conditional: $P_{x|y}(X \mid y) = P(X=x | Y=y) = \frac{P(X=x, Y= y}{P(Y=y)}= \frac{P_{x,y}(X,y)}{P_y(y)} = \frac{\text{joint}}{\text{marginal}}$
\end{enumerate}



\chapter{}
\section{Lecture 04} 
\subsection{Properties of condition expectation} 
\[E[X | X=x] = x \forall x\]


\[E[X|Y=y]\], is a number bu $E[X|Y]$ is a random variable. 

\[X \perp W, X \sim exp(\lambda), W \sim exp(\lambda)\]

\[f_x(x) = \lambda e^{-x} \text{for x} \ge 0\]

$Y= X + W$, condition on $Y=y$, $E[X|Y=y]$. We need $f_{x|y}(x|y)$

\[X| y + W = [y-X | y]\] 

This should have the same distributions. 


To show this, we need to get the distribution function. 

\[f_{x|y}(x|y) \frac{f_{x,y}(x,y)}{f_y(y)} = \frac{f_{x,w}(x, y-x}{f_y(y)} = \frac{f_x(x)f_w(y-x)}{f_y(y)} = \lambda e^{-\lambda x} \lambda e^{-\lambda(y-x)} = \lambda^2 e^{-\lambda y}\]
For $0\le x \le y$ 


We also need to get the marginal distributions. 

\[f_y(y) = \lambda^2 \int_{-\infty}^\infty f_{x,y}(x,y) dx = \lambda ^2 \int _0^y e^{-\lambda y}dx = \lambda ^2 ye^{-\lambda y}\]

\[f_{x|y}(x|y) = \frac{f_{x,y} (x,y)}{f_y(y)} \frac{\lambda ^2 e^{-\lambda y}}{y\lambda ^2 e^{\lambda y}} = \frac{1}{y}\]

	This is a uniform distribution from $0\le x\le y$. Therefore $E[X|Y=y] = \int_0^y (x \frac{1}{y})dx = \frac{y}{2}$ 



Instead, if we take $G(y) = E[X|Y]$, which is a random variable, we can get $E[E[X|Y]$. 

We can do this by $E[g(y)] = \int_{-\infty}^\infty g(x)D(x)dx$, which is the probability $g(X)$ multiplied by the density $d(x)$. 


\[E[E[X | Y]] =  \int E[X| Y=y] f_y(y) dy\]. 

It turns out that this becomes $E[X]$. 

\ex{}{$Y = \sum^N X_i, EX_i = \mu$, $N \perp X_i$'s, or N is independent of the $X_i$s.  

	\[E[Y] = E[\sum^N y_n] = E[E[\sum^N X_i | N]] = \sum_{\text{all } n} E[\sum^N X_i | N = n] p_n(n) = \sum_{\text{all} n} E[\sum m^n X_1] p_n(n) = \sum_{\text{all} n} n\mu p_n(n) \]
	\[=\mu \sum n p_n(n) = \mu EN\]
Where N is discrete, integer valued, nonnegative. 
$p_n(n)$ is the pmf for n 
}

\ex{Mean of a geometric}{

	\[X \sim geom(p)\]

Find EX. 

Because geometric dists are memoryless, we can condition on the first step. 

Let $I = I \{\text{1st trial is a success}\} \sim Bern(p)$ 

Compute $EX$
 
$$EX = E[ E[X|I]] = \sum_{\text{all i}} E[X | I = i) p(I=i) = E[X| I = 1] p(I=1) + E[X | I= 0 ] p(I=0)$$

$$ = 1p + (1+EX)(1-p) \implies EX = \frac{1}{p}$$ 

Basically saying, we either get the heads on this trial, or then we have to restart. 


}


\ex{Packet Transmission}{
Serve 1 packet per time slot. Each slot we have $A_i$ = \# arrivals in slot $i$. We assume i.i.d. and has $p_0, p_1, p_2$. 

We assume that the buffer is size 2. (We can only hold 2 packets in the buffer) 

Basically, for every time slot, we can only store up to 2 packets. Each timeslot, one packet leaves the buffer, and if there is too many packets, the incoming packet has to bounce. 

\[N = \text{length of busy period}\]
A busy is period is where there are arrivals to an empty system and ends when the system is emptied again. 

Need $E[N]$ 

First thing that we need to think about is conditioning. It would be helpful to know how many packets start the busy period, therefore we condition on $X = \# \text{packets that start the busy period (b.p)}$, where $X \in \{1, 2\}$. Thus we also need to know $P(X=1) = P(A = 1 | A=1 \text{or} = 2) = \frac{p_1}{p_1 + p_2} $ and $P(X=2) i = P(A =2 | \dots) = \frac{p_2}{p_1+p_2}$.  

\[E[N] = E[E[N|X]] = EN|X = 1] \frac{p_1}{p_1 + p_2} + E[N|X=2] \frac{p_2}{p_1+p_2} \]
}

\chapter{}
\section{Lecture 05} 

\subsection{Packet Transmission (cont)}

\ex{Packet Transmission (cont)}{

Packet transmission appends $ A_i $ i.i.d. $ p_0, p_1, p_2 $. 
$ N =  $ length of busy period. We still want $ E[N] $ 

\[X = \begin{cases} 1 \frac{p_1}{p_1+p_2} \\ 2 \frac{p_2}{p_1+p_2} \end{cases}\]
\[E[N] = E[E[N|X]] = E[N|X = 1] \frac{p_1}{p_1 + p_2} + E[N|X=2] \frac{p_2}{p_1+p_2} \]

We now need to compute $ E[N|X=n] $. 

For simplicity $ E[N_1] = E[N| X=1] $:

It would help to know how many packets we get at time 1

\[E[N_1] = E[E[N_1 | A]]\]

where $ A $  is the number of arrivals in the slot of the busy period. 

\begin{align*}
	E[N] &=  E[N_1 | A = 0]p_0 + E[N_1 | A=1]p_1  + E[N_1 | A=2]p_2\\ 
	&= 1p_0 + (1+E[N_1])p_1 + (1+E[N_2])p_2  \\
.\end{align*}

We can use the 	$ 1 + E[N_1] $ because of the i.i.d. and the memory less state. 

\[E[N_2] = E[N_2 | A=0]p_0 + E[N_2 | A=1]p_1 + E[N_2 | A=2]p_2\]

\[E[N_2] = (1+E[N_1])p_0 + (1+E[N_2])p_1 + (1+E[N_2]) p_2\]

The last term is because the buffer only has storage for 2 packets. If one goes out, then one gets replaced and the other gets sent away. 

We can thus solve for $ E[N_1] $ and $ E[N_2] $, which we can plug into the original equation. 

This is a very simple Markov Chain.
\vspace{10mm}


}

\subsection{Computing Variance:} 

$ \Var(x) $  in two methods

\begin{enumerate}
	\item Use $ \Var(x) = E[X^2] - E[X]^2 $, condition for $ E[X^2] + E[X] $ 
	\item Condition Var formula
\end{enumerate}

Example for method 1: 

\[X \sim geo(p)\]

\[E[X^2] = E[E[X^2]| I] \text{where} I  = \text{indicator of success } \sim \Bern(p) \]

\[E[X^2] = E[X^2 | I =1]p + E[X^2| I=0](1-p)\]
 \[ = 1p + (E[(X+1)^2])(1-p) = p + E[X^2 + 2x +1](1-p)\]
\[= p + (E[X^2] + E[2X] + 1 )(1-p)\]

Solve for $ E[X^2] = \frac{n-p}{p^2}$

\[\Var(X) = E[X^2] - (E[X]) ^2  = \frac{2-p}{p^2} - \left( \frac{1}{p} \right) ^2 = \frac{1-p}{p^2} \]


How do we get the method 2 formula? 

\proof{Conditional Variance Formula}{

	\begin{align*}
		[E[\Var(X|Y)] + \Var(E[X|Y]) &= E[E[X|Y ^2] - (E[X|Y])^2] + E[E[X|Y]^2] - E[E[E[X|Y]^2]]^2 \\
		&= E[E[X^2|Y]] - E[E[X|Y]]^2 + E[X^2] - E[X]^2 \\ 
		&=  \Var(X) \\
	.\end{align*}


}


\thm{$ X\perp Y \implies \Var(XY) = \Var(X)$}{
\[X\perp Y \implies \Var(XY) = E[Y^2]\Var(X)+ E[X]^2 \Var(Y) = E[X^2]\Var(Y) + E[Y]^2 \Var(X) \]

\begin{align*}
	\Var(XY) &= E[\Var(XY | Y)] + \Var(E[XY | Y]) \\ 
.\end{align*}

For the first term: 
\[\Var(XY | Y = y) = \Var(Xy | Y = y) = y^2\Var(X) \implies \Var(XY | Y) = Y^2 \Var(X)\]	
We can do this because $ X $  is independent of Y. 

\vspace{5mm}
For the second term:

\[E[XY | Y=y] = yE[X] \implies E[XY | Y] = YE[X] \]

Therefore, 
\[\Var(XY) = E[Y^2 \Var(X)] + \Var(YE[X]) = \Var(X)E[Y^2] + E[X]^2 \Var(Y)\]
}

\thm{}{
$ X_i $ is i.i.d. $ N $ is an integer valued random variable independent of the $ X_i $ s. Find the $ \Var(\sum_{}^{N} X_i) $


$$ \Var(\sum_{}^{N}X_i) = E[\Var(\sum_{}^{N}X_i | N)]  + \Var(E[\sum_{}^{N}X_i|N)] $$

For the first term: 

\[\Var(\sum_{}^{N}X_i | N=n) = \Var(\sum_{}^{n} X_n |N=n) = \sum_{}^{N}\Var(X | N=n) = \sum_{}^{n}\Var(X) = n\Var(X_i)\]

We proved the second term earlier. 
\[E[\Var(\sum_{}^{N}X_i)] = E[N\Var(X_i)] + \Var(NE[X_i]) = E[N]\Var(X_i) + \Var(N) E[X_i]^2\]

If $ N \perp X $ 

\[\Var(NX) = E[X^2]\Var(N) + E[N^2]\Var(X)\]

This one is like the catering example, where everyone orders the same thing. The other is if all the $ X_i $ are i.i.d., therefore ordering different things. Therefore the first is more variable than the second.

}

\chapter{}
\section{Lecture 06} 

\subsection{Recall:}

\[
I\{A\} \sim \Bern(P(A)) \implies E[I\{A\}] = P(A) 
\]

\[
EX = E[E[X \mid y]]
\]
Therefore,

\begin{align*}
	P(A) &= E[I\{A\}] \\
	&= E[E[I\{A \mid Y\}]] \\
	&=  E[P(A \mid y)] \\
	&=  \sum_{y}^{}P(A \mid Y=y)p_y(y) 
	\quad \text{or} \int_{-\infty}^{\infty }P(A \mid Y=y) f_y(y) dy\\
	&= \text{Total probability rule} \\
\end{align*}

These concepts are important for Markov chains! 


\ex{}{i.i.d. rolls of 2 fair dice. What the chance of sum =5 before sum =7


We could enumerate all the possibilities. 

\[
P(5 \text{on a single trial}) = \frac{4}{36}
\]

\[
P(7 \text{on a single trial} = \frac{6}{36}
\]

We can do the computation 2 ways. 

\[
	P(5 \text{ before } 7) = P(E)
\]

Conditioning!! 

Let $ X $  = sum on 1st roll. 
\begin{align*}
	P(E) &= E[P(E | X)] = P(E \mid X=5) P(X=5) + P(E \mid X=7)P(X=7) + P(E \mid X = \text{other} P(\text{other})) \\	
&= 1 \cdot \frac{4}{36} + 0 \frac{6}{36} + P(E)\frac{26}{36} \\
&= \frac{4}{10} = \frac{2}{5} \\
\end{align*}


Another way to do it: 

\[
P(E) = E[P(E \mid N)]
\]
Where $ N $ is the number of rolls until 5 or 7. $ N \sim geom(\frac{10}{36}) $ 

\begin{align*}
	P(E) &= \sum_{n=1}^{\infty }P(\text{sum to 5 on nth roll for 1st time} \mid N = \text{first 5 or 7}) P(N=n)\\
&= \sum_{n=1}^{\infty }P(\text{sum to 5 on roll n} \mid \text{sum to 5 or 7 on roll n })P(N=n)	 \\
&= \sum_{n=1}^{\infty } \frac{\frac{4}{36}}{\frac{4}{36} + \frac{6}{36}}P(N=n)  \\
&=  \sum_{n=1}^{\infty } \frac{4}{10}P(N=n) \\
&= \frac{2}{5} \sum_{n}^{}P(N=n) \\
&= \frac{2}{5} \cdot 1
\end{align*}

because of i.i.d., it doesn't matter whether if its the first time or not 


We show here that $ E \perp N $. 
}

\subsection{Markov chains}
A stochastic process is a sequence of random variables. 

\[
\{X_n\} = \text{ stochastic process } = \text{ sequence of R.Vs}
\]

Simplest sequence: $ X_n $ i.i.d. This has no memory and is stateless. 

A Markov chain only considers the most recent thing that happened

MC = markov chain, DTMC = discrete time markov chain. 

\dfn{Discrete Time Markov Chain}{
A markov chain is defined as 

\[
	X_{n+1} \mid X_n, X_{n-1}, \dots X_0  \perp X_{n-1}, \dots, X_n
\]
\[
	P(X_{n+1} \mid X_n =, X_{n-1} =, \dots) = P(X_{n+1} | X_n) = P_{ij}
\]
In other words, only the previous state matters. 

For us, time homogeneous, N doesn't matter
}


\[
	P = ||P_{ij}||
\]

This is a transition matrix, where all elements are nonnegative, and the row sum = 1. From state $ i $ to state $ j $.  


\ex{Gilbert Model of transmission}{


	G is good State, B is Bad State

	We could either stay in each node, or to move between good and bad. 

Transition Matrix: 
\[	P = \begin{bmatrix}
		&1-p &p \\
		&q &1-q
	\end{bmatrix} \]

}
\ex{
4 state markov chain: 
}{
\[
X \in \{GG, GB, BG, BB\}
\]

\[
P(G \mid X) = p_1, p_2, p_3, p_4
\]

Going from $ GG $ to $ GG $, the probability is $ p_1 $. going from $ GG $ to $ GB $, that would be $ 1-p_1 $. 

From $ GB $ to $ BG $, the probability is $ p_2 $. Going from $ GB $ to $ BB $, the probability $ 1-p_2 $ 

etc. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%% ADD MARKOV DIAGRAM %%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
}

\ex{Birth Death Markov Chain}{
Infinite number of states, but only 2 transmissions. You can either go up and down, or stay in the same state. see textbook for more.
}

\ex{Inventory Model}{

$ D_i $ i.i.d. daily demand. If inventory $ \le s$ at day end, order up to $ S $ to start next day. $ (s, S) $ policy. 

State = start of day inventory level

\vspace{5mm} 
States are from $ \{s+1, s+2, i, S-1, S\} $

To stay at state $ i $ , $ P(D_i = 0) $. to up to $ S $, $ P(D_i \ge i-s) $. To go down to $ i-1 $, $ P(D_i = 1) $, etc.  
}

\chapter{}
\section{Lecture X: 25 February 2025}

$ \{X\} $ irreducible $ \iff $ there is only one class for all states ($ i \leftrightarrow j \forall i,j \in S $ 

\dfn{Period}{
\[
d(i) = \text{gcd}\{N \ge 1 : P_{ii}^{n} > 0\}
\]
$ i $ has period $ d = d(i): $

\[
P_{ii}^{n} > 0 \implies \text{d divides n}
\]
And all states have the same period. 


\vspace{5mm}
n is not a multiple of $ d \implies P_{ii}^{n} = 0 $, thus aperiodic : $ d(i) = 1 $ 
}
\ex{Period of 2}{
\begin{figure}[H]
	\centering
	\begin{tikzpicture}[->,>=stealth',shorten >=2pt , line width =0.5pt, node distance =2cm]
	\node [circle, draw] (i-1)  {i-1};
	\node [circle, draw] (i) [right=of i-1] {i};
	\node [circle, draw] (i+1) [right=of i] {i+1};
	\path (i-1) edge [bend left] node [above] {$p$} (i);
	\path (i) edge [bend left] node [below] {$q$} (i-1);
	\path (i+1) edge [bend left] node [below] {$p$} (i);
\path (i) edge [bend left]  node [above] {$p$} (i+1);
	\end{tikzpicture}
	\caption{period of 2}
	\label{fig:period of 2 ex}
\end{figure}

there are more nodes past $ i+1 $ with same structure in a birth death cycle. Therefore, you can only get to even nodes or odd nodes depending on $ i $.  
}

\subsection{Stationary Distribution}
$ X $: $ X_{n} \sim \pi $, where $ \pi $ is a row vector for all the probabilities. $ \pi $ is the stationary distribution (equilibrium/invariant) $ n $, if $ X_{n} \sim \pi \implies X_{n+1} \sim \pi  $  


\ex{Gilbert Model}{

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[->,>=stealth',shorten >=2pt , line width =0.5pt, node distance =2cm]
	\node [circle, draw] (G)  {G};
	\node [circle, draw] (B) [right=of G] {B};
	\path (G) edge [bend left] node [above] {$.1$} (B);
	\path (B) edge [bend left] node [below] {$.2$} (G);
	\path (G) edge [loop left] node [left] {$.9$} (G);
	\path (B) edge [loop right] node [right] {$.8$} (B);
	\end{tikzpicture}
	\caption{Gilbert model}
	\label{fig:}
\end{figure}

Our intuition states that the stationary dist $ n $ should be $ \pi = (\frac{2}{3}, \frac{1}{3}) $, because we are staying i$ .\frac{.2}{.3} $ of the time into G. 

\vspace{5mm}
\[
	\pi_{G} = P(\text{end up in G)}) = P(\text{start in G and stay or start in B and move})
\]

\begin{align*}
	&= \pi_{g}(1-p) + \pi_{B}(q) \\
	&= p\pi_{g} = q\pi_{B} \\
\end{align*}

If we solve this for B as well, and use the fact that $ \pi = 1 $. We get $ \pi = (\pi_{g}, \pi_{b}) = (\frac{q}{p+q}, \frac{p}{p+q} $)

Plugging in $ p=.1, q=.2 $, we get our original answer. 

\[
\pi_{G} = \text{rate of transitions into or out of G}
\]
In other words, detailed balance of $ \pi_{G}(1-p) = \pi_{B}q $ 
}


\ex{Packet Transmissions}{
Packet example: (G/D/1/2) (Incoming, Outgoing (Deterministic), server \#, memory) 

\[
A_{i} \sim (p_{0}, p_{1}, p_{2})
\]
\begin{figure}[H]
	\centering
	\begin{tikzpicture}[->,>=stealth',shorten >=2pt , line width =0.5pt, node distance =2cm]
	\node [circle, draw] (0) {0};
	\node [circle, draw] (1) [right=of 0] {1};
	\node [circle, draw] (2) [right=of 1] {2};
	\path (0) edge [loop left] node [left] {$p_{0}$} (0);
	\path (0) edge [bend left] node [above] {$p_{1}$} (1);
	\path (0) edge [bend left=60] node [above] {$p_{2}$} (2);
	\path (1) edge [bend left] node [below] {$p_{0}$} (0);
	\path (1) edge [loop below] node [below] {$p_{1}$} (1);
	\path (1) edge [bend left] node [above] {$p_{2}$} (2);
	\path (2) edge [bend left] node [below] {$p_{0}$} (1);
	\path (2) edge [loop right] node [right] {$p_{1} + p_{2}$} (2);
	\end{tikzpicture}
	\caption{packet transmission}
	\label{fig:Packet Transmission}
\end{figure}

What is the stationary dist $ n $?

\[
\pi_{0} = \pi_{0}p_{0} + \pi_{1}p_{0} + \pi_{2}\cdot 0
\]
\[
\pi_{1} = \pi_{0}p_{1} + \pi_{1}p_{1} + \pi_{2}p_{0}
\]
\[
\pi_{2} = \pi_{0}p_{2} + p_{1}p_{2} + \pi_{2}(p_{1} + p_{2})
\]
\[
\pi_{1} + \pi_{2} + \pi_{3} = 1
\]

This results in $ \pi = (\frac{p_{0}}{1-p_{1}} + \frac{p_{0}(1-p_{0})}{1-p_{0}}, \frac{p_{2}}{1-p_{1}} )$ 


}

\thm{Stationary Distribution Computation}{

	If $ \pi $ satisfies 
	\begin{enumerate}
	\item $ \pi_{j} = \sum_{i\in S}^{} \pi_{i}P_{ij} \forall j \iff \pi =\pi P$ 
		\item $ \sum_{i\in S}^{} \pi_{1} = 1 $ 
	\end{enumerate}

	Then $ \pi $ is a stationary distribution



}
\proof{
Suppose $ X_{0} \sim \pi $, We need to show that $ X_{1} \sim \pi $

For $ P\{X_{1} = j \} = \pi_{j} $ 

	\[
		P\{X_{1} = j \} = E[P(X_{1 = j \mid X_{0}})] = \sum_{i \in S}^{} P(X_{1} = j \mid X_0 = i)P(X_{0} = i) = \sum_{i \in S}^{P_{ij} \cdot \pi_{i}}
	\]
}



	\dfn{Steady State}{

		The steady state distribuion of $ \{X\} $, when it exists, is defined for all j by 

		\[
		\lim_{n\rightarrow \infty } P(X_{n} = j \mid X_{0} = i) = \lim_{n \rightarrow \infty }P_{ij}^{n} = \text{steady state prob of j}
		\]
	}


\dfn{Long run time average distribution}
{
	the long run time distribution of $ \{X\} $, if it exists, is defined for all j by 
	\[
	\lim_{n\rightarrow \infty }\frac{1}{n} \sum_{m=0}^{n}I\{X_{m} = j \mid X_{0} = i\} = \text{l.r. prop time in j}
	\]

This is weaker than the steady state, basically like an average. counting the number of times we are in state $ n $, resulting in the indicators. 
}

\thm{}{For irreducible periodic, positive recurrent MC, 

\[
P_{ij}^{n} \rightarrow \pi_{j}, \quad \text{where $ \pi $ is the stationary distribution}
\]
}

\thm{}{For irreducible positive recurrent MC, long run proportion time in j = $ \pi_{j} $, where $ \pi $ is the stationary distribution. 

}a

\chapter{}
\section{Lecture 09} 

\subsection{Stationary Distribution} 
\ex{Gilbert Model}{
For the gilbert Model: 
\begin{tikzpicture}[->, >=stealth, node distance=3cm, every state/.style={thick, fill=gray!10}]

  % States
  \node[circle] (G) {(G)};
  \node[circle, right of=G] (B) {(B)};

  % Transitions
  \path (G) edge[loop left] node[left] {$1-p$} (G)
        (G) edge[bend left] node[above] {$p$} (B)
        (B) edge[bend left] node[below] {$q$} (G)
        (B) edge[loop right] node[right] {$1-q$} (B);

\end{tikzpicture}
If $ p=.1 $ and $ q=.2 $ 
\[
\pi_{G} = \pi_{B}(q) + \pi_{G}(1-p)
\]
\[
\pi_{B} = \pi_{B}(1-q) + \pi_{G}p
\]

AND 

\[
\pi_{G} + \pi_{B} = 1
\]

\[
\pi = (\pi_{G}, \pi_{B}) = (\frac{q}{p+q}, \frac{p}{p+q})
\]

\begin{align*}
	P_{11}^{1} &=  .9 \\
	P_{11}^{2} &=  .9^{2} + (.1)(.2) = .82 \\
P_{11}^{3} &= .781 \\
\vdots\\
P_{11}^{n} &=  \frac{2}{3}
\end{align*}

This converges because the chain is aperiodic, irreducible, finite state. 
}
\ex{Packet Transmission Revisited}{
For the Packet Transmission \ref{fig:Packet Transmission}: G/D/1/2

\[
A_{i} \sim (p_0, p_1, p_2) 
\]
if $ p_{0} = p_{1} = \frac{1}{4} $ 
We will get $ \pi = \left( \frac{1}{12}, \frac{1}{4}, \frac{2}{3} \right) $ 

The $ P(\text{busy}) $ in the long run, it will be $ \frac{11}{12} $.

\vspace{5mm}
What is the loss rate? $ E[\text{\# of packets lost/slot}] $ 

\begin{align*}
	E[\text{\# of packets lost}] &= \pi_{2}p_2(1)	\\
	&= \frac{2}{3} \cdot \frac{1}{2} \\
	&= \frac{1}{3} \\
\end{align*}

On average, once every 3 time slots, we lose a packet. 

\vspace{5mm}
What proportion of packets are lost? We need the arrival rate = \# packets coming per slot. 

\begin{align*}
	E[A_{i}] &= 0p_0 + 1p_1 + 2p_2 \\ 
	&= \frac{1}{4} + 2 \cdot \frac{1}{2}  = \frac{5}{4}  \\
\end{align*}

Thus $ \frac{1}{3} \div \frac{5}{4} = \frac{4}{15} $. This is the number that are lost per slot divided by number of packets arriving per slot. Thus, about $ .272 $ of the packets are lost. 

\vspace{5mm}
What is the effective service rate? What is the through put of the system? 

\begin{align*}
\text{throughput}	&= \text{proportion served $ \times $ arrival rate} \\
&= (1-\text{prop lost)} \cdot \text{arrival rate}  \\
&= (1-\frac{4}{15}) \cdot \frac{5}{4} \\
&= \frac{11}{12} \\
&= P(\text{busy}) \cdot 1 \text{served per slot} \\
\end{align*}
}
\ex{Google Search Rankings}{
What order to put pages in? Develop a giant markov chain, with the websites being a state and the probability to visit that state. 


\begin{figure}[H]
	\centering
	\begin{tikzpicture}[->,>=stealth',shorten >=2pt , line width =0.5pt, node distance =2cm]
	\node [circle, draw] (page i)  {page i};
	\path (-2, -1) edge node {} (page i); 
	\path (-2, 1) edge node {} (page i); 
	\path (-2, 0) edge node {} (page i); 
\path (page i) edge  node  {} (2,0);

	\end{tikzpicture}
	\caption{Google Page Ranking}
	\label{fig:Google Page Ranking}
\end{figure}

So multiple sites pointing into page i, and then also they have a weighted arrows going out. 
}

\subsection{Special Cases} 
$ p $ is doubly stochastic. 

For a doubly stochastic markov chain, the rows and the columns sum up to 1. 

In this case, all the states are equally likely, or a uniform distribution

\[
\pi = (\frac{1}{n}, \frac{1}{n}, \dots
\]

Cut sets: $ \forall  \text{cut} $ 
If we partition our states, the rate across one way is equal to the rate across the other way . This is especially helpful in birth death processes (ex. Gilbert Model) 

If we cut the Gilbert Model down the middle, 
\[
\pi_{G}p = \pi_{B}q
\]


\begin{figure}[H]
	\centering
	\begin{tikzpicture}[->,>=stealth',shorten >=2pt , line width =0.5pt, node distance =2cm]
	\node [circle, draw] (i)  {i};
	\node [circle, draw] (i-1) [left=of i] {i-1};
	\node [circle, draw] (i+1) [right=of i] {i+1};

	\path (i-1) edge [bend left] node [above] {$p_{i-1}$} (i);
	\path (i) edge [bend left] node [below] {$1-p_{i}$} (i-1);
	\path (i) edge [bend left] node [above] {$p_{1}$} (i+1);
	\path (i+1) edge [bend left] node [below] {$1-p_{i+1}$} (i);
	\end{tikzpicture}
	\caption{Birth Death}
	\label{fig:Birth Death}
\end{figure}


If we cut this between i and i+1, 
\begin{align*}
	\pi &= \pi P \\
	\pi_{i}& = \pi_{i-1}p_{i-1} + pi_{i+1}p_{i+1}\\
	\pi_{i+1} &= \pi_{i}(\frac{p_{1}}{p_{i+1}})\\
\end{align*}

\vspace{5mm}
\[
m_{jj} = E[\text{\# of transitions, back to j}] = E[T_{j}]
\]
Thm: For an irreducible positive recurrent MC, $ \pi_{j} = \frac{1}{E[T_{j}]} = \frac{1}{m_{jj}} \forall j $ 

For the gilbert Model, $ p=.1, q=.2 $. If we want the $ E[T_{g}] $ or the time to get from good to good. 

\begin{align*}
E[T_{g}] &=  \frac{1}{\pi_{g}} \\
&= \frac{1}{\frac{2}{3}} \\ 
&= \frac{3}{2} \\ 
\end{align*}

The long way: 
\begin{align*}
	E[T_{g}] &= .9(1 + 0) + .1(1+ E[T_{BG}]) \\
	&= 1 + .1E[T_{BG}] \\
	&= 1 + .1(5) \\
	&= \frac{3}{2} \\
\end{align*}

$ T_{BG} $ has a geometric distribution, $ \sim geometric(.2) $. Therefore $ E[T_{BG}] = 5 $ 

\chapter{}
\section{Lecture 11: 4 March 2026}

Suppose $ \{X_{n}\} $ has stationary distribuion $ \pi $. 

\begin{itemize}
	\item Let $ X $ bc r.v $ X \sim \pi $ 
	\item Let $ r $ be a function $ S \to \mathbb{R} $ s.t. $ r(x) $ be the cost when state is $ x $. 
\end{itemize}

\dfn{LRAC}{The long run average cost (LRAC) is:
	\[
	LRAC = \lim_{n\to \infty } \sum_{n=0}^{N}\frac{r(x_{n}}{N}
	\]


}

\thm{When LRAC, exists, LRAC = $ E[r(x)] $}{
\begin{align*}
	LRAC &=  E[r(x)] \\
	&= \sum_{i \in S}^{}r(i) \pi_{i} \\
\end{align*}

A sufficient condition for LRAC existence: 

\begin{itemize}
	\item $ \{X\} $ irreducible, positive recurrent and $ r $ bounded or linear.
\end{itemize}
}

\cor{When $ r(x) $ is linear, $ r(x) = a +bx  $ then LRAC = $ a+bE[X], X \sim\pi $  }

\cor{If you pay \$1 per unit time in state $ j $, 0 otherwise. Then $ r(x) = I\{X=j\} $ and LRAC = $ \pi_{j} $ } 

\ex{}{
Inventory: $ s=0, S=3 $ , 0 lead time, no back orders. $ X_{n} $ = inventory at start of day. Daily demand $ D_{i} $ i.i.d. and $ D_{i} \sim (.2, .4, .1, .3) $

\begin{figure}[H]
	\centering
	\begin{tikzpicture}[->,>=stealth',shorten >=2pt , line width =0.5pt, node distance =2cm]
	\node [circle, draw] (1)  {1};
	\node [circle, draw] (2) [right=of 1] {2};
	\node [circle, draw] (3) [right=of 2] {3};

	\path (1) edge [loop left] node [left] {$.2$} (1);
	\path (1) edge [bend left=60] node [above] {$.8$} (3);
	\path (2) edge [loop below] node [below] {$.2$} (2);
	\path (2) edge [bend left] node [below] {$.4$} (1);
	\path (2) edge [bend left] node [above] {$.4$} (3);
	\path (3) edge [loop right] node [right] {$.5$} (3);
	\path (3) edge [bend left] node [below] {$.4$} (2);
	\path (3) edge [bend left=60] node [below] {$.1$} (1);
	\end{tikzpicture}
	\caption{Inventory Model}
	\label{fig:Inventory Model}
\end{figure}

\nt{Because we only order when we get to 0 and only order at the end of the day, we can only go from $ 1 \to 1 $ and $ 1 \to 3 $. Similarity for 3. We can get to 3 with $ D_{i}=0  $ or $ D_{i}=3 $ because we can order at the end of the day.} 


\[
	\mathbb{P} = \begin{bmatrix}
		.2 & 0 &.3 \\
		.4 &.2 &.4 \\
		.1 &.4 &.5
	\end{bmatrix}
\]


We can solve the balance equations $ \pi = \pi P $. 

\begin{align*}
	\pi_{1} &= .2\pi_{1} + .4\pi_{2} + .1 \pi_{3}\\
		&\vdots \\
\pi &=  1 \\
\end{align*}

\[
\pi = \left (\frac{1}{5}, \frac{4}{15}, \frac{8}{15}\right)
\]
\begin{align*}
E[X] &= 1 \cdot \frac{1}{5} + 2 \cdot \frac{4}{15} + 3 \cdot \frac{8}{15} \\
&= \frac{7}{3} \\
&= 2.33 \\
\end{align*}

So the long run average holding cost if costs is $ \$2 $ unit/day: 

\begin{align*}
	E[r(x)] &= 2\cdot 1 \pi_{1} + 2 \cdot 2 \pi_{2} + 2 \cdot 3 \pi_{3} \\
	&= 2E[X] \\
	&= 4.67 \\
\end{align*}

So on average, we are paying $ \$4.67 $ per day in holding costs.

}

\ex{Extension of 7.1.1 with \href{fig:Inventory Model}{Inventory Model}}{
What is LRAC w/ holding cost (h.c.) $ \$2 $ per unit per day, ordering cost $ \$3 $ per order + stockout cost = \$4 unit stocked out.

\vspace{5mm}
For ordering cost, need reorder rate (RR) = long run rate of transitions triggering an order.  Or in other words, long run proportion of days where demand is $ \ge $ supply .

\begin{align*}
	RR &=\pi_{1}(P(D_{i} \ge 1) + \pi_{2}(P(D_{i}\ge 2)) + \pi_{3}(P(D_{i} \ge 3)) \\
	&= .8 \cdot \frac{1}{5} + .4 \cdot \frac{4}{15} + .3 \cdot \frac{8}{15} \\
	&\approx .43  \\
\end{align*}

So the l.r average reorder cost is \$3 per order(.43 orders per day) = \$1.29 per day. 

\vspace{5mm}
Stock out cost: 

We need long run average number of units stocked out. (SR) j

\begin{align*}
	SR &= \pi_{1}(1 P(D_{i}=2) + 2 P(D_{i} = 3)) + \pi_{2}\left( 2 P(D_{i}=3) \right) \\
	&= \pi_{1}(.1 + 2 \cdot .3) + \pi_{2}(.3) \\
	&= .22 \\
\end{align*}

We have on average 1 stock out every 4 days. Therefore LR average stock out cost is (\$4 per unit)(.22 units/day) = \$ .88 per day.

\vspace{5mm}
Now we have the total costs, we can start the analyze the policy and optimize for min costs.

}

\ex{$ S_{ij}  = P(\text{ever get to j} \mid X_{0} = i)$
}{
$ S_{ij}  = P(\text{ever get to j} \mid X_{0} = i)$

\vspace{5mm}
If $ j $ is recurrent and $ j \in R $. $ i $ is also recurrent. 

\begin{item}
\item 	If $ i, j $ is in the same recurrent class, then $ f_{ij} = 1 $ 
\item If $ i \in R' \neq R $, then $ f_{ij} = 0 $, because recurrent classes are closed.
\end{item}

\vspace{5mm}
If $ i $ transient and $ j$ recurrent , or $ j $ transient then we change to make absorbing so then it makes it recurrent. 

\[
T = \text{set of transient states}
\]

Let $ j\in R \forall i \in T$:

Then 

\begin{align*}
	f_{ij} = \sum_{k \in T}P_{ik} f_{kj} + \sum_{k\in R}^{P_{ik}}  0
\end{align*}
Basically, if we go to a node in T, or go to the rccurent state, or go into another reccurent class. 
}


\chapter{}

\subsection{Lecture 12: 9 March 2026}


\chapter{}
\section{Lecture 13: 11 March 2026} 

\subsubsection{Exponential Distribution}

\[
f(x) = \lambda e^{-\lambda x }, x \ge 0 
\]

$ \lambda $ is the rate, and $ SD(X) = \frac{1}{\lambda} $ and $ \Var(X) = \frac{1}{\lambda^{2}} $ 
Property: 
\begin{enumerate}
	\item $ X \sim exp(\lambda) \implies X $ memoryless (see Data 140 notes) k
	\item $ P(X_1 < X_2) = P(\min(X_1, X_2)) = \frac{\lambda_{1}}{\lambda_{1} + \lambda _{2}}$ 
	\proof{
			\begin{align*}
				P(X_1 < X_2) &= E[P(X_1 < X_2 \mid X_1)] \\
				&= \int_{0}^{\infty } P(X_1 < X_2 \mid X=x)\lambda e^{-\lambda x} dx \\
				&=  \int_{0}^{\infty} P(X_2 > X_1) \lambda_{1}e^{-\lambda_{1}x}dx   \\
				&= \int_{0}^{\infty } e^{-\lambda_{2}x} \lambda_{1}e^{-\lambda_{1}x} dx \\
				&= \frac{\lambda_{1}}{\lambda_{1} + \lambda_{2}} \\
			\end{align*}
		}

	\item $ \min(X_{i}) = exp(\sum_{}^{}\lambda_{i} $ 

		\proof{ Show that $ P(\min(X_{1}) > x)  \le e^{-\sum_{}^{}\lambda x}$. This is a tail distribution.

			\begin{align*}
				P(\min(X_{i} \ge x ) &= \prod P(X_{i} \ge x) \\
				&= \prod e^{-\lambda_{i} x} \\
				&= e^{-\sum_{}^{}\lambda_{i}x} \\
			\end{align*}

		}
	\end{enumerate}


Thus, $ EX_{t} = EX = \frac{1}{\lambda} $, while $ E[X \mid X > t] = t + \frac{1}{\lambda} $ 

\ex{Post Office}{


Post office serves based on $ exp(\lambda) $. Brown, Smith and Jones walk in, and they are served i.i.d.. What is the probability that Smith is the last to leave? 

Jones and Brown are initally being served. 

Suppose that Jones finishes first, and Smith comes in.
They are actually equally likely to leave. $ \frac{1}{2} $ 


\vspace{5mm}
What is the probability that Jones is the last to leave? 

Then it would be $ \frac{1}{4} $ by the memoryless property. 


}

\ex{Failure rates}{

if $ X_{1} \sim ex(1), X_2 \sim ex(2)$

then $ P(X_{1} < X_{2}) = \frac{1}{3} $ and the $ \min(X_1, X_2) \sim exp(3) $ 
}

\end{document}
